Different centralities in multilayer stream graphs

### 2019-07-29 tex/article/main.tex
ajout des pdf autres

\section{Introduction}
%
From the time Euler, in 1750 stated the bridges problem of Köninsgberg, the notion of graphs has been widely exploited to model and sort a lot of real-life and theoretical problems.

@@

A {\em multilayer graph}~\cite{mlkiv,mldd} is a set $M=(V_M,E_M,V,{\cal L})$.\\
${\cal L}$ is the {\em structure}: it is a finite set of $d$ different sets, named {\em aspects}: ${\cal L} = L_1, \dots, L_d$. Each aspect $L_i$ contains elements $l_i^1,\dots l_i^{n_i}$ which are named {\em elementary layers}. A {\em layer} $\alpha$ is then a combination of elementary layers from each aspect: $\alpha \in L=L_1\times \dots \times L_k$. \\
$V$ is the set of nodes and each of them can be present or not on the different layers. A node on a layer will be called a {\em node-layer}. The set of nodes-layers is $V_M \subseteq V \times L$. Each node-layer can be linked to another with undirected ($E_M\subseteq V_M \otimes V_M$)or directed ($E_M \subseteq V_M \times V_M$) edges.

@@

We built a object containing both the temporal and structural aspect, we will now explain how to do the reverse operation, how we can extract some stream graphs, multilayer graphs and graphs from a multilayer stream graph.

@@

\subsection{Density}
%
In a classical graph $G=(V,E)$, the density is the rate of connections: $\delta(G)= \frac{|E|}{|V|^2}$ if the graph is directed, and  $\delta(G)= \frac{2|E|}{|V|(|V|-1)}$ if it's not directed. We will focus on the undirected multilayer stream graphs.

\begin{definition}[Density of a multilayer stream graph]	
	Nous appelons  $C \in T \times V_M\times V_M$ l'ensemble des liens autorisés dans le \stgm{}. Si certaines liens sont \og automatiques\fg{} ou \og sous-entendus \fg{} comme par exemple les arêtes de couplage, nous ne les comptons pas dans $C$ non plus pour que le cas où la densité soit nulle corresponde au cas où on ne peut pas retirer de lien.

	Let's call $C \in T \times V_M\times V_M$ the set of ``possible links'' in the multilayer stream graph. If some links are implied (the links between the nodes-layers from the same nodes), they are not in $C$.

@@

In our case, we have nodes, layers, node-layers, and links. This means that the measures will get more complex and more diverse.

@@

Two types of groups of layers are preponderant in multilayer graphs, the superimposed layers and the juxtaposed layers, and we will define centralities for the two types of layers.

The {\bf superimposed layers}: in this group of layers, nodes can appear on all layers. It is the case for layers describing to diverse types of relationships. If two different node-layers are from different nodes and different superimposed layers, we don't have a link between those two node-layers. A relevant way to study those layers will be to compare them.

The {\bf juxtaposed layers} forms a partition of the nodes. For instance, they can model people from different classes in the same school. A relevant way to deal with those layers will be to study the interactions between them.

\subsubsection{Centrality of times}
%
\subsubsection{Centrality of nodes}
%
\subsubsection{Experimental centralities: random walk}
%

>>>>

\section{Introduction}
%
From the time, in 1750, that stated the bridges problem of Köninsgberg, the notion of graphs has been widely exploited to model and sort a lot of real-life and theoretical problems.

@@

A {\em multilayer graph}~\cite{mlkiv,mldd} is a set $M=(V_M,E_M,V,{\cal L})$.\\
${\cal L}$ is the {\em structure}: it is a finite set of $d$ different sets, named {\em aspects}: ${\cal L} = L_1, \dots, L_d$. Each aspect $L_i$ contains elements $l_i^1,\dots l_i^{n_i}$ which are named {\em elementary layers}. A {\em layer} $\alpha$ is then a combination of elementary layers from each aspect: $\alpha \in L=L_1\times \dots \times L_k$. \\
$V$ is the set of nodes and each of them can be present or not on the different layers. A node on a layer is called a {\em node-layer}. The set of nodes-layers is $V_M \subseteq V \times L$. Each node-layer can be linked to another with undirected ($E_M\subseteq V_M \otimes V_M$)or directed ($E_M \subseteq V_M \times V_M$) edges.

@@

We built a object containing both the temporal and structural aspect, we now explain how to do the reverse operation, how we can extract some stream graphs, multilayer graphs and graphs from a multilayer stream graph.

@@

\subsection{Density}
%
In a classical graph $G=(V,E)$, the density is the rate of connections: $\delta(G)= \frac{|E|}{|V|^2}$ if the graph is directed, and  $\delta(G)= \frac{2|E|}{|V|(|V|-1)}$ if it's not directed. We focus on the undirected multilayer stream graphs.

\begin{definition}[Density of a multilayer stream graph]	Let's call $C \in T \times V_M\times V_M$ the set of ``possible links'' in the multilayer stream graph. If some links are implied (the links between the nodes-layers from the same nodes), they are not in $C$.

@@

In our case, we have nodes, layers, node-layers, and links. This means that the measures become more complex and more diverse.
%
\subsubsection{Centrality of nodes and the paths}
%
In classical graph, the centrality of the nodes describe how much we will encounter a node while browsing a graph. They are different way to browse a graph: the random walk, the contamination models, the shortest path, etc. In the stream graph, the notions are more numerous: we can compute random walks, but with different laws of probabilities of transition, same for the contamination. We can look for the shortest paths (with the less transitions possible) but the fastest paths (the ones that take less time), the foremost paths (the ones that arrives the earliest starting from a certain time), and we can combine those notions.

@@

Two types of groups of layers are preponderant in multilayer graphs, the superimposed layers and the juxtaposed layers, and we define centralities for the two types of layers.

The {\bf superimposed layers}: in this group of layers, nodes can appear on all layers. It is the case for layers describing to diverse types of relationships. If two different node-layers are from different nodes and different superimposed layers, we don't have a link between those two node-layers. A relevant way to study those layers is be to compare them.

The {\bf juxtaposed layers} forms a partition of the nodes. For instance, they can model people from different classes in the same school. A relevant way to deal with those layers is to study the interactions between them.

\subsubsection{Centrality of times}
%

\subsubsection{Experimental centralities: random walk}
%
